---
title: "03 CSI online typing: Manual answer classification and participant exclusion"
author: "Kirsten Stark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r load_packages}
rm(list = ls())

library(tidyr)
library(dplyr)
```


# Load data and answercodes

```{r}
options( "encoding" = "UTF-8" )

# input 
input = "data_long.csv"

# load data
df <- read.csv(here::here("data", "transient_data_files", input))

# load answer codes
answercodes <- read.csv(here::here("data", "supplementary_info", "answercodes.csv"),sep = ";")
```

# Check and classify participants' answer behavior
Assign answer codes to participants' typed naming

```{r}
df$answercode <- NA
for(i in 1:nrow(answercodes)) {
  df$answercode[toupper(df$item) == toupper(answercodes$item[i]) &
                  toupper(df$word) == toupper(answercodes$word[i])] <- 
    answercodes$answercode[i]
}
df$answercode <- ifelse(is.na(df$word), "isna", df$answercode)
```

Have all answers been classified? 

```{r}
sum(is.na(df$answercode)) == 0
```

Overview of answer codes (naming mistakes) - BEFORE participant exclusion:

```{r}
table(df$answercode)
```

- *Almost correct* naming are either synonyms or naming with only minor or corrected typos where the word is still recognizable and the initial letter is correct.  
- Naming was classified as *backspace_space_enter* when the they started by typing the backspace, space, enter, or caps lock key, and then typed some word.  
- *Correct* are all word entrys that are identical to the naming participants were familiarized with.  
- *First_letter_incorrect* are words/synonymes that are later typed (almost) correct, but where the first letter was wrong, even if it was later on corrected. The entries suggest that participants sometimes continued typing the last word if they hadn't finished to do so.  
- *isna* when no typed entry was recorded.  
- *semantic_relation* when another semantically related word was typed that was not a synonym.  
- *shift_start* when participants started typing the word with pressing the shift button. 
- *unrelated_other* are either completely unrelated words, nonwords, single letter entries, or only keys like backspace, space, enter.  
  
**For the main analysis, we consider correct and almost correct entries as correct:**

```{r}
df$correct <- NA
df$correct <- ifelse(df$answercode == "correct" | df$answercode == "almostcorrect", 1, NA)

sum(df$answercode == "correct") + sum(df$answercode == "almostcorrect") ==
  sum(df$correct, na.rm = TRUE)
```

## How many trial onset times are considered as invalid?

```{r}
# Amount and percentage of NA or excluded trials
(sumna <- sum(is.na(df$correct)))
(percentagena <- sum(is.na(df$correct))/nrow(df))
```

In a total of `r sumna` trials participants did not enter anything or entered a chunks of characters that were not considered as correct based on the preregistered criteria. This is `r round(percentagena,2)` % of all trials.   
  
*Trials with missing values per participant:*

```{r}
for (i in 1:length(unique(df$type))) {
  print(paste(unique(df$type)[i],"data collection: ", sep = " "))
  print("Amount of trials without any entry for timing.01")
  print(as.data.frame(table(
    df$subject[df$type == unique(df$type)[i]],
    is.na(df$timing.01[df$type == unique(df$type)[i]]))) %>% 
      filter(Var2 == TRUE) %>%
  dplyr::rename(subject = Var1, totaltrials = Var2, NA_trials = Freq) %>%
  mutate(totaltrials = 160) %>% 
    mutate(percentage_NA = NA_trials/totaltrials))
}
```

*Amount of trials classified as correct per participant:*

```{r}
for (i in 1:length(unique(df$type))) {
  print(paste(unique(df$type)[i],"data collection: ", sep = " "))  
  print("Amount of trials classified as correct 
        (correct, correct with typos, synonymes")
  print(as.data.frame(table(df$subject[df$type == unique(df$type)[i]],
                      !is.na(df$correct[df$type == unique(df$type)[i]]))) %>%
    filter(Var2 == TRUE) %>%
    dplyr::rename(subject = Var1, totaltrials = Var2, correct = Freq) %>%
      mutate(totaltrials = 160) %>% 
      mutate(percentagecorrect = correct/totaltrials))
}
```

# Participant exclusion
## Naming performance 

As preregistered, participants with an accuracy below 80 % (either because of inaccurate naming or because of no naming) will be excluded and replaced for the main analyses:

```{r}
for (i in 1:length(unique(df$type))) {
  print(paste(unique(df$type)[i],"data collection: ", sep = " "))  
   print("Participants with performance below 80 %")
   print(as.data.frame(table(df$subject[df$type == unique(df$type)[i]],
                      !is.na(df$correct[df$type == unique(df$type)[i]]))) %>%
    filter(Var2 == TRUE) %>%
    dplyr::rename(subject = Var1, totaltrials = Var2, correct = Freq) %>%
     mutate(totaltrials = 160) %>% 
     mutate(percentagecorrect = correct/totaltrials) %>%
     filter(percentagecorrect < .80))
}
```

**In the main data collection, participants with subject no. 3 and 9 will be excluded from the data analyses.**

## Comments

```{r}
#table(unique(df$comments[df$comments != "NA"]))
as.data.frame(table(df$type, df$subject, df$comments)) %>% filter(Freq != 0)
```

We will exclude one participant, the participant with the subject no. 20, who indicated that she has been living in England for several years and is writing usually in English and therefore at first only remembered the English word of the items and is unfamiliar with typing Umlaute. Her answer behavior also indicated that she typed the English words several times.

```{r}
unique(df$comments[df$subject == 20 & df$type == "main"])
```

## Attention checks
*1) Item vs. non-item*  
We had two item- vs. non-item attention checks

```{r}
## Item vs. non-item
# CH01_01 (Taube), CH01_02 (Apfel), CH02_01 (Luftballon) and CH02_02 (Biene) are items and 2 should appear in the dataframe,
# CH02_03 (Radio), CH02_04 (Sparschwein), CH02_03 (Laptop) and CH02_04 (Wattestäbchen) are non-items and 1 should appear in the dataframe
# --> As subjects had to select two items, 
# controlling the selected items is enough
for(i in 1:length(unique(df$type))) {
  print(as.data.frame(df %>% filter(type == unique(df$type)[i]) %>%
  select(type, subject, CH01_01, CH01_02, CH02_01, CH02_02)) %>% 
  filter( CH01_01 == 1 | CH01_02 == 1 | CH02_01 == 1 | CH02_02 == 1)%>%
  unique())
}

```

6 participants of the main data collection (subject no. 6, 11, 14, 22, 24, 29) made one mistake in the item vs. non-item task after the familiarization phase. However, all performed perfectly in the item vs. non-item task after the main task. 

*2) Cheating*

```{r}
## Did participants cheat
# CH03 = 1 - yes, I worked through it till the end,
# CH03 = 2 - no, I stopped or cheated midway
# CH03 = -9 - no answer
for(i in 1:length(unique(df$type))) {
  print(as.data.frame(df %>% filter(type == unique(df$type)[i]) %>%
  select(type, subject, CH03)) %>% 
  filter(CH03 != 1)%>%
  unique())
}
```

One participant in the main data collection indicated that they stopped working on the task or cheated midway. However, in the comments, that participant indicated that they only indicated this because they had to open the door during the familiarization phase and therefore stopped for about 2 min. As there was no timeout in the familiarization, this shouldn't have affected the data much. Therefore, we do not exclude this participant.

```{r}
unique(df$comments[df$subject == 19 & df$type == "main"])
```

*3) Mother tongue*

```{r}
table(df$language)
table(df$language.test)
```



*4) Keyboard*

```{r}
table(df$keyboard_type)
table(df$KB02_01)
table(df$KB03_01)
table(df$KB03_02)
table(df$KB03_03)
```

The keyboard screening worked well for all participants.

### Exclude participants and export final data frame

```{r}
df2 <- df %>% 
  filter(!((df$subject == 3 | df$subject == 9 | 
              df$subject == 20) & df$type == "main"))# %>% 

df2$subject <-  ifelse(df2$subject == 1 & df2$type == "replacement", 9, df2$subject)
df2$subject <- ifelse(df2$subject == 2 & df2$type == "replacement", 20, df2$subject)
df2$subject <- ifelse(df2$subject == 3 & df2$type == "replacement", 3, df2$subject)
table(df2$subject)

write.csv(df2, here::here("data", "transient_data_files", "data_long_final.csv"))
```

