---
title: '05 CSI online typing: Plotting and analysis'
author: "Kirsten Stark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages
```{r load_packages}
library(dplyr)
library(tidyr)
library(lme4)
library(lmerTest)
library(Rmisc)
library(Cairo)
#library(strengejacke)
library(ggplot2)
library(sjPlot)

options(scipen=999)

rm(list = ls())
options( "encoding" = "UTF-8" )
set.seed(99)
```


## Load and preprocess data

```{r load_data}
# input 
#input = "data_long_final.csv"
input = "data_long_anonymous.csv"
classification_type = "automatic" # select "manual" or "automatic"

# load data
df <- read.csv(here::here("data", input))

```

Check amount of participants and trials

```{r checks}
# no. of participants: 
length(unique(df$subject))

# no. of trials is 160 per participant? 
nrow(df) == 160 * length(unique(df$subject))
```

Factorize columns

```{r transform_variables}
# factorize columns
is.numeric(df$timing.01)
df$PosOr <- as.factor(df$PosOr)
df$subject <- as.factor(df$subject)
```

# Select correct classification column

```{r classification_type}
if(classification_type == "automatic") {
  df$answercode <- df$answer_auto_jaro
  df$correct <- df$correct_auto_jaro
} else if(classification_type == "manual") {
  df$answercode <- df$answercode
  df$correct <- df$correct_manual
} else {
  print("Select a correct type!")
}
```

```{r answercode_table}
as.data.frame(table(df$correct, df$answercode)) %>% filter(Freq != 0) 
as.data.frame(table(df$correct, df$answercode)) %>% filter(Freq != 0) %>% 
  mutate(Percentage = case_when(Var1 == 1 ~ Freq/sum(df$correct == 1),
                                Var1 == 0 ~ Freq/sum(df$correct == 0)))
```

```{r table_correctness}
# raw
table(df$correct)
# in percent
round(table(df$correct)/nrow(df)*100,2)

## How many correct/incorrect non-filler trials?
table(df$correct[df$category != "Filler"])
```

Show amount of incorrect trials per ordinal position (excluding fillers):

```{r}
## How many correct/incorrect non-filler trials per ordinal position?
table(df$PosOr[df$category != "Filler" & df$correct == 0], 
      df$correct[df$category != "Filler" & df$correct == 0])

```

Drop incorrect trials: 

```{r}
df <- df %>% filter(df$correct == 1)
```


# Plotting

Make plots suitable for APA format, font sizes can be adjusted

```{r apatheme}
apatheme <- theme_bw()+
  theme(plot.title=element_text(family="Arial",size=22,hjust = .5),
        panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        panel.border=element_blank(),axis.line=element_line(),
        text=element_text(family="Arial",size=16))
```

### Descriptives

```{r descriptives}
(means_final<- df %>% 
   filter(category != "Filler") %>% 
   Rmisc::summarySEwithin(.,"timing.01",idvar = "subject",
                          withinvars = "PosOr", na.rm = T))

# Export as word file
library(flextable)
huxt_word <- huxtable::huxtable(means_final)
huxt_word <- huxtable::set_number_format(huxt_word, round(2))
huxtable::quick_docx(huxt_word, 
                     file = here::here("results", "tables",
                                       "CSI_online_typing_RT_summary.docx"), 
                                       open = FALSE)

```

### RTs by ordinal position
Line graph (only correct trials, without fillers)

```{r plot_rt}
(plot_rt <- means_final %>% 
    ggplot(., aes(x=PosOr, y=timing.01)) +
    geom_point(size = 2)+
    stat_summary(fun=mean,  geom="line", size = 0.5, group = 1, linetype = "dashed") +
    geom_errorbar(aes(ymin=timing.01-se, ymax=timing.01+se), width =.1) +
    apatheme+
    scale_y_continuous(limits = c(1120, 1340), breaks =seq(1120,1340, by = 20)) +
                       #breaks = c(1100, 1150, 1200, 1250, 1300, 1350)) + 
    labs(x="Ordinal Position ",y ="RT (ms)") + #+
  # annotate(geom="text", x=1.5, y=1330, label="n = 30", 
  #         color="black", size = 8))
    theme(
    axis.title.y = element_text(margin = margin(0,10,0,0)),
    axis.title.x = element_text(margin = margin(10,0,0,0))))

filename <- "CSI_online_typing_plot_rt.pdf"
ggsave(plot_rt, filename = 
         here::here("results", "figures", filename),
       width = 18, height = 13, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("results", "figures", filename))

```


Normalized boxplot 

```{r boxplot_rt}
means_subject <- df %>% 
   filter(category != "Filler") %>% 
   summarySEwithin(.,"timing.01",withinvars = c("subject","PosOr"))
(means_subject <- means_subject %>%
  group_by(subject) %>%
  dplyr::mutate(timing.01_norm = timing.01 - first(timing.01)))

(boxplot <- 
  ggplot() + 
  
  ## boxplot
  geom_boxplot(data=means_subject, aes(x = PosOr,y =timing.01_norm),
               colour = "grey", width = 0.3,fatten = 1)+
  ### individual means
  geom_jitter(data=means_subject, aes(x = PosOr,y =timing.01_norm),
              position = position_dodge(0.6),
              shape=19,color = "dark grey", size=2)+
  ### group means
  stat_summary(data=means_subject, aes(x = PosOr,y =timing.01_norm),
               fun=mean, geom="point",colour = "black", shape=18, size=5)+
  ### line
  stat_summary(data=means_subject, aes(x = PosOr,y =timing.01_norm),
               fun=mean, geom="line",colour = "black", linetype = "longdash", group = 1)+
  
  ## other stuff
  #scale_y_continuous(breaks = seq(600, 1300, by = 50))+
  labs(x="Ordinal Position",y ="Normalized RTs (ms)")+
  apatheme +
  theme(
    axis.title.y = element_text(margin = margin(0,10,0,0)),
    axis.title.x = element_text(margin = margin(10,0,0,0))) +
  coord_equal(ratio = 1/100))

filename <- "CSI_online_typing_boxplot.pdf"
ggsave(boxplot, filename = 
         here::here("results", "figures", filename),
       width = 13, height = 18, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("results", "figures", filename))

```

### Export plot grid

```{r plot_grid}
cowplot::plot_grid(plot_rt, boxplot,
          nrow = 1, labels = c("A", "B"), label_fontfamily = "Arial") %>%
  ggsave(filename = here::here("results", "figures",
                               "CSI_online_typing_RTs_and_normalized_RTs"),
         width = 18, height = 13, units = "cm", dpi = 300, 
         device = cairo_pdf)
#embedFonts(file = here::here("results", "figures", "CSI_online_typing_RTs_and_normalized_RTs"))

```

### ... with fillers for control

```{r plot_rt_fillers}

(plot_rt_fillers <- df %>% 
    mutate(kind = case_when(category == "Filler" ~"Filler",
                          category != "Filler" ~"Experimental")) %>%
    ggplot(., aes(x=PosOr, y=timing.01, group=kind, color=kind)) +
    stat_summary(fun=mean,  geom="point", size = 2)+
    stat_summary(fun=mean,  geom="line", size = 1) +
    apatheme+
    labs(x="Ordinal Position ",y ="RT (ms)", color = "Trial type")+
  annotate(geom="text", x=1.5, y=1350, label="n = 30", 
           color="black", size = 8))

filename <- "CSI_online_typing_plot_rt_with_fillers.pdf"
ggsave(plot_rt_fillers, filename = 
         here::here("results", "figures", filename),
       width = 18, height = 13, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("results", "figures", filename))
```
  
### Plot by subcategory

```{r plot_rt_by_cat}
(plot_rt_by_cat <- df %>%  
   filter(category != "Filler") %>% 
    ggplot(., aes(x=PosOr, y=timing.01)) +
    stat_summary(fun=mean,  geom="point", size = 2)+
    stat_summary(fun=mean,  geom="line", size = 1) +
    facet_wrap(~category) +
    apatheme+
    labs(x="Ordinal Position ",y ="RT (ms)"))

filename <- "CSI_online_typing_plot_rt_by_category.pdf"
ggsave(plot_rt_by_cat, filename = 
         here::here("results", "figures", filename),
       width = 18, height = 19, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("results", "figures", filename))

```

### Plot by subject

```{r plot_by_subject}
(plot_rt_by_subject <- df %>% 
   filter(category != "Filler") %>% 
    ggplot(., aes(x=PosOr, y=timing.01)) +
    stat_summary(fun=mean,  geom="point", size = 2) +  
    stat_summary(fun=mean,  geom="line", size = 1) +  
    facet_wrap(~subject) +
    apatheme+
    labs(x="Ordinal Position ",y ="RT (ms)"))

filename <- "CSI_online_typing_plot_rt_by_subject.pdf"
ggsave(plot_rt_by_subject, filename = 
         here::here("results", "figures", filename),
       width = 18, height = 19, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("results", "figures", filename))

```

### Control: Plot RTs accross the experiment
All trials correct trials

```{r plot_RTs_all}
(plot_RTs_all <- ggplot(data=df, aes(x=trial, y=timing.01)) +
  stat_summary(fun=mean,  geom="point", size = 2)+
  stat_summary(fun=mean,  geom="line", size = 1) +
  apatheme+
  labs(x="Trial ",y ="RT (ms)")+
  annotate(geom="text", x=20, y=1570, label="n = 30", 
           color="black", size = 8))

filename <- "CSI_online_typing_plot_rts_across_experiment.pdf"
ggsave(plot_RTs_all, filename = 
         here::here("results", "figures", filename),
       width = 18, height = 13, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("results", "figures", filename))
```

Correct non-filler trials only:

```{r plot_RTs_correct}
(plot_RTs_correct <- df %>% 
   filter(category != "Filler") %>% 
    ggplot(., aes(x=trial, y=timing.01)) +
    stat_summary(fun=mean,  geom="point", size = 2)+
    stat_summary(fun=mean,  geom="line", size = 1) +
    apatheme+
    labs(x="Trial ",y ="RT ms")+
    annotate(geom="text", x=20, y=1570, label="n = 30", 
             color="black", size = 8))

filename <- "CSI_online_typing_plot_rts_across_experiment_correct_experimental_trials.pdf"
ggsave(plot_RTs_correct, filename = 
         here::here("results", "figures", filename),
       width = 18, height = 13, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("results", "figures", filename))
```

# Check distribution of data
Are the data normally distributed or does a gamma distribution fit the data better?  
*Subset data to correct trials only and exclude fillers*

```{r final_data}
df_valid <- df %>% filter(category != "Filler") %>% 
  filter(correct == 1) %>% droplevels()
```

*Center predictor variable*

```{r}
df_valid$PosOr.cont <- scale(as.numeric(as.character(df_valid$PosOr)),
                                        center = T, scale = F)
# table(df_valid$PosOr.cont)
# mean(df_valid$PosOr.cont)
```

*Histogram of the reaction time data*

```{r RT_hist}
hist(df_valid$timing.01)
```

*Check fit of normal vs gamma distribution in histograms, q-q-plots and using objective criteria:*  
1) Fit normal and gamma distributions to the reaction time data

```{r load_fitdistrplus}
library(fitdistrplus)
```

```{r fit.normal}
fit.normal<- fitdist(df_valid$timing.01, distr = "norm", method = "mle")
summary(fit.normal)
#plot(fit.normal)
```

```{r fit.gamma}
fit.gamma <- fitdist(df_valid$timing.01, distr = "gamma", method = "mle")
summary(fit.gamma)
#plot(fit.gamma)
```

2) Compare the fit of the two distributions  
Visually compare fit of both distributions in histogram

```{r density-comparison}
denscomp(list(fit.gamma, fit.normal))
```

Visually compare fit of both distributions in Q-Q-plots

```{r q-q-comparison}
qqcomp(list(fit.gamma, fit.normal))
```

Compare information criteria

```{r fit_criteria-comparison}
gofstat(list(fit.gamma, fit.normal),
        fitnames = c("Gamma", "Normal"))
```

**Conclusion:** Both the visual inspection and the objective criteria suggest that a gamma distribution fits the data better (although not that well). Therefore, we fit a Gamma distribution in a GLMM with the continuous predictor ordinal position (Pos.cont), the factorial predictor (experiment), and their interaction. We compute the maximal random effects structure.


# Inferential analyses: GLMM (Gamma distribution) with ordinal position as a continuous predictor


```{r GLMM_cont}
m1 <- glmer(timing.01 ~ PosOr.cont + 
               (PosOr.cont|subject) +(PosOr.cont|category),
             data = df_valid, 
            family =Gamma(link ="identity"), 
            control=glmerControl(optimizer = "bobyqa"))
summary(m1)

# save model output
tab_model(m1,transform = NULL,
          show.re.var = F, show.stat = T,show.r2 = F,show.icc = F,
          title = "GLMM (Gamma distribution) with continuous predictor",
          pred.labels = c("(Intercept)", "Ordinal Position"),
          dv.labels = "Typing Onset Latency",
          #string.pred = "",
          string.stat = "t-Value",
          file = here::here("results", "tables", "CSI_online_typing_glmm_cont.html"))
```



# ---------------------------------------------
# Appendix
Line graph for each participant: 

```{r plot_by_size}
modeloutput <- coef(m1)$subject
(means_final_subject <- df_valid %>% 
   summarySEwithin(.,"timing.01",withinvars = c("subject","PosOr")))
(means_final<- df_valid %>%  
   Rmisc::summarySEwithin(.,"timing.01",idvar = "subject",
                          withinvars = "PosOr", na.rm = T))
for(i in 1:nrow(means_final_subject)) {
  means_final_subject$grandmean[i] <- means_final$timing.01[means_final$PosOr == means_final_subject$PosOr[i]] - means_final$timing.01[means_final$PosOr== 1]
  means_final_subject$normalizedRT[i] <- means_final_subject$timing.01[i] -
    means_final_subject$timing.01[means_final_subject$subject == means_final_subject$subject[i] & means_final_subject$PosOr == 1]
  # prepare for ordering
  means_final_subject$effect[i] <- modeloutput$PosOr.cont[means_final_subject$subject[i]] 
}
#means_final_subject$order <-order(means_final_subject$effect)-1
#means_final_subject$order <- means_final_subject$order %/% 5
means_final_subject <- means_final_subject[order(desc(means_final_subject$effect)),] 
means_final_subject$effect <- as.factor(round(means_final_subject$effect, 2))
means_final_subject$effect <- factor(means_final_subject$effect, levels=rev(levels(means_final_subject$effect )))

(plot_rt <- means_final_subject %>% 
    ggplot(., aes(x=PosOr,y=normalizedRT, na.rm=T)) +
    #geom_point(aes(x=PosOr,y=grandmean), size = 0.5, color = 'gray')+
    geom_line(aes(x=PosOr,y=grandmean), group = 1,size = 0.8, color = 'dark gray', linetype = 3)+
    geom_point(size =1, color = 'black') +
    geom_line(group = 1,size = 0.5, color = 'black', linetype = "dashed") +
    #geom_errorbar(aes(ymin=timing.01-se, ymax=timing.01+se), width =.1) +
    apatheme+
    scale_y_continuous(limits = c(-200, 600), 
                       breaks = c(-200, 0,200,400,600)) + 
    labs(x="Ordinal Position",y ="Normalized RTs (ms)") +
    facet_wrap(means_final_subject$effect))

filename <- "CSI_online_typing_effect_by_participant.pdf"
ggsave(plot_rt, filename = 
         here::here("results", "figures", filename),
       width = 13, height = 18, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("results", "figures", filename))

```
