---
title: "04 CSI online typing: Descriptives"
author: "Kirsten Stark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r load_packages}
library(dplyr)
library(tidyr)

rm(list = ls())
```


## Load and preprocess data

```{r load_data}
options( "encoding" = "UTF-8" )

# input
input <- "data_long_final.csv"

# load data
df <- read.csv(here::here("data", input), sep = ",",  na = "")
```

## Duration of the experiment

```{r duration}
print("Outlier-corrected duration (provided by soscisurvey)")
mean(df$time_wo_outlier)/60 # 21, 37 min = 21 min 22sec
sd(df$time_wo_outlier)/60
range(df$time_wo_outlier)/60 # 0.00000 30.21667
duration <- df %>% dplyr::select(starts_with("TIME")) %>% 
  dplyr::select(!"time_wo_outlier") %>% dplyr::select(!"TIME_RSI") 
print("Outlier-corrected duration (provided by soscisurvey)")
duration$sum = duration %>% rowSums(na.rm = TRUE)
mean(duration$sum)/60
sd(duration$sum)/60
range(duration$sum)/60
```


## Description of participants

Gender: 

```{r gender}
df <- df %>% mutate(gender_char = case_when(gender == 1 ~ "female", 
                                     gender == 2 ~ "male"))
table(df$gender_char)/160 # 1 = female, 2 = male, 3 = diverse

print("percentage female:")
sum(df$gender == 1)/nrow(df)
```

Age: 

```{r age}
print("mean:"); mean(df$age)
print("sd:"); sd(df$age)
print("range:"); range(df$age)
```

Handedness: 

```{r handedness}
# 1 = left handed, 2 = right handed, 3 = ambidexter/both
df <- df %>% mutate(handedness_char = case_when(handedness == 1 ~ "left-handed", 
                                     handedness == 2 ~ "right-handed"))
table(df$handedness_char)/160 

print("percentage right-handed:")
sum(df$handedness == 2)/nrow(df)
```

Fingers used for typing: 

```{r finders_used}
# 1 = 1, 2 = 2, 3 = 3, 4 = 4, 5 = 5, 6 = don't know
print("left hand: "); table(df$fingers_l)/160
print("right hand: "); table(df$fingers_r)/160

df$fingers_l <- ifelse(df$fingers_l == 6, NA, df$fingers_l)
df$fingers_r <- ifelse(df$fingers_r == 6, NA, df$fingers_r)

print("Average number of fingers used (both hands combined):") 
mean(df$fingers_l+df$fingers_r, na.rm = T); sd(df$fingers_l+df$fingers_r, na.rm = T)
```

Mother tongue (experiment was restricted to native German speakers): This seems to have worked

```{r}
table(df$language) # 1 = yes (mother tongue is German), 2 = no
table(df$language.test) # 1 = der, 2 = die, 3 = das (das is correct)
```


## Average typing speed and accuracy

Subset all typing test columns and select one row per particpant only 

```{r preparation typing speed}
source("automatic_preprocessing_functions.R")
dat <- df %>% filter(trial == 1) %>%
    dplyr::select(c("subject", starts_with("TT")))
```

```{r read_solutions}
test1 <- "Der Fuchs ist nah mit Hund und Wolf verwandt. Die Tiere sehen Hunden auch recht ähnlich. Füchse haben jedoch einen längeren Körper und etwas kürzere Beine."
print("Actual length of characters in test1 (including spaces):")
nchar(test1)

test2 <- "Der Schwanz eines Fuchses ist fast halb so lang wie das ganze Tier. Das Fell ist meist rot oder braun, der Bauch weiß und die Schwanzspitze hell oder dunkel."
print("Actual length of characters in test2 (including spaces):")
nchar(test2)

test3 <- "Typisch für den Fuchs sind zudem seine aufgestellten Ohren, die ihm auf der Jagd behilflich sind. Füchse hören Mäuse aus hundert Meter Entfernung quieken."
print("Actual length of characters in test3 (including spaces):")
nchar(test3)
```

### Accuracy

Accurary similar as in e.g., Crump (2003), Crump & Logan (2010), Pinet, Dubarry, & Alario (2016): Percentage of 5-character words containing no error (neither a backspace nor a typographical error). The data from the three text subsections will be collapsed.  
**!!!!!!!! CAVE: UNFORTUNATELY, THERE WAS AN ERROR IN THE EXPERIMENT AND ONLY THE FIRST 23 SINGLE WORDS ARE RECORDED FOR EACH TEXT. THE TOTAL TYPING TIME AND THE ENTIRE TEXT RECORDING ARE STILL CORRECT FOR THE WHOLE TEXT!!!!!!!!**  
**7 participants failed to disable caps lock and wrote everything in uppercase. Here, the accuracy is estimated in uppercase writing** 

```{r fiveletterword_accuracy}

# Create wordbased df
fivecharwords <- data.frame(
  text = c(rep(1, times=26), rep(2, times = 29), rep(3, times = 24)),
  word_no = c(seq(1,26, by=1), seq(1,29,by=1), seq(1,24, by=1)), 
  word = c("Der","Fuchs", "ist", "nah", "mit", "Hund", "und", "Wolf",
              "verwandt", "Die", "Tiere", "sehen", "Hunden", "auch",  "recht",
              "ähnlich", "Füchse", "haben", "jedoch", "einen", "längeren",
              "Körper", "und", "etwas", "kürzere", "Beine",
            
              "Der", "Schwanz", "eines", "Fuchses", "ist", "fast", "halb", 
              "so", "lang", "wie", "das", "ganze", "Tier", "Das", "Fell", "ist",
              "meist", "rot", "oder", "braun", "der", "Bauch", "weiß", "und",
              "die", "Schwanzspitze", "hell", "oder", "dunkel",
            
              "Typisch", "für", "den", "Fuchs", "sind", "zudem", "seine", 
              "aufgestellten", "Ohren", "die", "ihm", "auf", "der", "Jagd",
              "behilflich", "sind", "Füchse", "hören", "Mäuse", "aus",
              "hundert", "Meter", "Entfernung", "quieken"))
fivecharwords <- fivecharwords %>% 
                  mutate(nchar=nchar(word)) %>%
                  filter(nchar == 5)%>% 
                  filter(word_no <= 23)

# copy actual typed word into the dataframe
for(i in 1:nrow(fivecharwords)) {
  if(fivecharwords$word_no[i] <= 23) {
      for(j in 1:length(dat$subject)) {
    if(fivecharwords$word_no[i] < 10) {
      eval(parse(text=paste0("fivecharwords$subject", j, "[", i, "]<-", 
                             "dat$TT0", fivecharwords$text[i]*3, "_0",
                             fivecharwords$word_no[i],"[",j,"]")))
    } else {
      eval(parse(text=paste0("fivecharwords$subject", j, "[", i, "]<-", 
                             "dat$TT0", fivecharwords$text[i]*3, "_",
                             fivecharwords$word_no[i],"[",j,"]")))    
    }
  }
  }
}

## Manually correct if word count is messed up, and set to NA if word number would have been > 23
# Necessary updates of word numbers in text 1
fivecharwords$subject18[1] <- dat$TT03_03[18] #; dat$TT03_02[18]; dat$TT03_03[18]
fivecharwords$subject19[1] <- dat$TT03_03[19] #; dat$TT03_02[19]; dat$TT03_03[19]; 
fivecharwords$subject1[2] <- dat$TT03_12[1] #; dat$TT03_11[1]; dat$TT03_12[1]; 
fivecharwords$subject3[2] <- dat$TT03_12[3] #; dat$TT03_11[3]; dat$TT03_12[3]; 
fivecharwords$subject6[2] <- dat$TT03_13[6] #;dat$TT03_11[6]; dat$TT03_13[6]; 
# subject 7 deleted the first nine words and then disabled capslock
fivecharwords$subject7[2] <- dat$TT03_20[7]  #;dat$TT03_11[7]; dat$TT03_20[7]; 
fivecharwords$subject8[2] <- dat$TT03_14[8] #; dat$TT03_11[8]; dat$TT03_14[8]; 
fivecharwords$subject12[2] <- dat$TT03_12[12] #; dat$TT03_11[12]; dat$TT03_12[12]; 
fivecharwords$subject16[2] <- dat$TT03_13[16] #;dat$TT03_11[16]; dat$TT03_13[16]; 
fivecharwords$subject18[2] <- dat$TT03_12[18] #; 
fivecharwords$subject19[2] <- dat$TT03_12[19] #;dat$TT03_11[19]; dat$TT03_12[19]; 
fivecharwords$subject21[2] <- dat$TT03_10[21]#;dat$TT03_11[21]; dat$TT03_10[21]; 
fivecharwords$subject25[2] <- dat$TT03_12[25]#;dat$TT03_11[25]; dat$TT03_12[25]; 
fivecharwords$subject26[2] <- dat$TT03_14[26]#;dat$TT03_11[26]; dat$TT03_14[26]; 
fivecharwords$subject28[2] <- dat$TT03_13[28]#;dat$TT03_11[28]; dat$TT03_13[28]; 

fivecharwords$subject1[3] <- dat$TT03_13[1]#;dat$TT03_12[1]; dat$TT03_13[1]; 
fivecharwords$subject3[3] <- dat$TT03_13[3]#;dat$TT03_12[3]; dat$TT03_13[3]; 
fivecharwords$subject6[3] <- dat$TT03_14[6]#;dat$TT03_12[6]; dat$TT03_14[6]; 
fivecharwords$subject7[3] <- dat$TT03_21[7] #;dat$TT03_12[7]; dat$TT03_21[7]; 
fivecharwords$subject8[3] <- dat$TT03_15[8]#;dat$TT03_12[8]; dat$TT03_15[8]; 
fivecharwords$subject12[3] <- dat$TT03_13[12]#;dat$TT03_12[12]; dat$TT03_13[12]; 
fivecharwords$subject13[3] <- dat$TT03_13[12]#;dat$TT03_12[13]; dat$TT03_13[13]; 
fivecharwords$subject16[3] <- dat$TT03_14[16]#;dat$TT03_12[16]; dat$TT03_14[16];
fivecharwords$subject18[3] <- dat$TT03_13[18]#;dat$TT03_12[18]; dat$TT03_13[18]; 
fivecharwords$subject19[3] <- dat$TT03_13[19]#;dat$TT03_12[19]; dat$TT03_13[19]; 
fivecharwords$subject21[3] <- dat$TT03_11[21]#;dat$TT03_12[21]; dat$TT03_11[21];
fivecharwords$subject25[3] <- dat$TT03_13[25]#;dat$TT03_12[25]; dat$TT03_13[25]; 
fivecharwords$subject26[3] <- dat$TT03_15[26]#;dat$TT03_12[26]; dat$TT03_15[26]; 
fivecharwords$subject28[3] <- dat$TT03_14[28]#;dat$TT03_12[28]; dat$TT03_14[28]; 

fivecharwords$subject1[4] <- dat$TT03_16[1]#;dat$TT03_15[1]; dat$TT03_16[1]; 
fivecharwords$subject3[4] <- dat$TT03_16[3]#;dat$TT03_15[3]; dat$TT03_16[3]; 
fivecharwords$subject6[4] <- dat$TT03_17[6]#;dat$TT03_15[6]; dat$TT03_17[6]; 
fivecharwords$subject8[4] <- dat$TT03_18[8]#;dat$TT03_15[8]; dat$TT03_18[8];
fivecharwords$subject12[4] <- dat$TT03_18[12]#;dat$TT03_15[12]; dat$TT03_18[12]; 
fivecharwords$subject13[4] <- dat$TT03_18[12]#;dat$TT03_15[13]; dat$TT03_18[13]; 
fivecharwords$subject16[4] <- dat$TT03_17[16]#;dat$TT03_15[16]; dat$TT03_17[16]; 
fivecharwords$subject18[4] <- dat$TT03_16[18]#;dat$TT03_15[18]; dat$TT03_16[18];
fivecharwords$subject19[4] <- dat$TT03_16[19]#;dat$TT03_15[19]; dat$TT03_16[19]; 
fivecharwords$subject21[4] <- dat$TT03_14[21]#;dat$TT03_15[21]; dat$TT03_14[21]; 
fivecharwords$subject26[4] <- dat$TT03_18[26]#;dat$TT03_15[26]; dat$TT03_18[26]; 
fivecharwords$subject28[4] <- dat$TT03_17[28]#;dat$TT03_15[28]; dat$TT03_17[28]; 
fivecharwords$subject7[4] <- NA
fivecharwords$subject25[4] <- NA

fivecharwords$subject1[5] <- dat$TT03_19[1]#;dat$TT03_18[1]; dat$TT03_19[1]; 
fivecharwords$subject3[5] <- dat$TT03_19[3]#;dat$TT03_18[3]; dat$TT03_19[3]; 
fivecharwords$subject6[5] <- dat$TT03_20[6]#;dat$TT03_18[6]; dat$TT03_20[6]; 
fivecharwords$subject8[5] <- dat$TT03_21[8]#;dat$TT03_18[8]; dat$TT03_21[8]; 
fivecharwords$subject12[5] <- dat$TT03_22[12]#;dat$TT03_18[12]; dat$TT03_22[12]; 
fivecharwords$subject13[5] <- dat$TT03_22[12]#;dat$TT03_18[13]; dat$TT03_22[13]; 
fivecharwords$subject16[5] <- dat$TT03_20[16]#;dat$TT03_18[16]; dat$TT03_20[16]; 
fivecharwords$subject18[5] <- dat$TT03_19[18]#;dat$TT03_18[18]; dat$TT03_19[18]; 
fivecharwords$subject19[5] <- dat$TT03_19[19]#;dat$TT03_18[19]; dat$TT03_19[19]; 
fivecharwords$subject23[5] <- dat$TT03_19[23]#;dat$TT03_18[23]; dat$TT03_19[23]; 
fivecharwords$subject26[5] <- dat$TT03_22[26]#;dat$TT03_18[26]; dat$TT03_22[26]; 
fivecharwords$subject28[5] <- dat$TT03_21[28]#;dat$TT03_18[28]; dat$TT03_21[28]; 
fivecharwords$subject7[5] <- NA
fivecharwords$subject25[5] <- NA

fivecharwords$subject1[6] <- dat$TT03_21[1]#;dat$TT03_20[1]; dat$TT03_21[1]; 
fivecharwords$subject3[6] <- dat$TT03_21[3]#;dat$TT03_20[3]; dat$TT03_21[3]; 
fivecharwords$subject6[6] <- dat$TT03_22[6]#;dat$TT03_20[6]; dat$TT03_22[6]; 
fivecharwords$subject8[6] <- dat$TT03_23[8]#;dat$TT03_20[8]; dat$TT03_23[8]; 
fivecharwords$subject16[6] <- dat$TT03_22[16]#;dat$TT03_20[16]; dat$TT03_22[16]; 
fivecharwords$subject18[6] <- dat$TT03_21[18]#;dat$TT03_20[18]; dat$TT03_21[18]; 
fivecharwords$subject19[6] <- dat$TT03_21[19]#;dat$TT03_20[19]; dat$TT03_21[19]; 
fivecharwords$subject23[6] <- dat$TT03_21[23]#;dat$TT03_20[23]; dat$TT03_21[23]; 
fivecharwords$subject7[6] <- NA
fivecharwords$subject25[6] <- NA
fivecharwords$subject12[6] <- NA
fivecharwords$subject13[6] <- NA
fivecharwords$subject26[6] <- NA
fivecharwords$subject28[6] <- NA

# Necessary updates of word numbers in text 2
fivecharwords$subject7[7] <- dat$TT06_04[7]#;dat$TT06_03[7]; dat$TT06_04[7]; 
fivecharwords$subject26[7] <- dat$TT06_05[26]#;dat$TT06_03[26]; dat$TT06_05[26]; 

fivecharwords$subject2[8] <- dat$TT06_13[2]#;dat$TT06_12[2]; dat$TT06_13[2]; 
fivecharwords$subject7[8] <- dat$TT06_14[7]#;dat$TT06_12[7]; dat$TT06_14[7]; 
fivecharwords$subject12[8] <- dat$TT06_13[12]#;dat$TT06_12[12]; dat$TT06_13[12]; 
fivecharwords$subject13[8] <- dat$TT06_13[13]#;dat$TT06_12[13]; dat$TT06_13[13]; 
fivecharwords$subject17[8] <- dat$TT06_14[17]#;dat$TT06_12[17]; dat$TT06_14[17]; 
fivecharwords$subject19[8] <- dat$TT06_13[19]#;dat$TT06_12[19]; dat$TT06_13[19]; 
fivecharwords$subject25[8] <- dat$TT06_13[25]#;dat$TT06_12[25]; dat$TT06_13[25]; 
fivecharwords$subject26[8] <- dat$TT06_15[26]#;dat$TT06_12[26]; dat$TT06_15[26]; 
fivecharwords$subject28[8] <- dat$TT06_13[28]#;dat$TT06_12[28]; dat$TT06_13[28]; 

fivecharwords$subject2[9] <- dat$TT06_18[2]#;dat$TT06_17[2]; dat$TT06_18[2]; 
fivecharwords$subject7[9] <- dat$TT06_18[7]#;dat$TT06_17[7]; dat$TT06_18[7]; 
fivecharwords$subject8[9] <- dat$TT06_20[8]#;dat$TT06_17[8]; dat$TT06_20[8]; 
fivecharwords$subject12[9] <- dat$TT06_18[12]#;dat$TT06_17[12]; dat$TT06_18[12]; 
fivecharwords$subject13[9] <- dat$TT06_18[13]#;dat$TT06_17[13]; dat$TT06_18[13]; 
fivecharwords$subject17[9] <- dat$TT06_22[17]#;dat$TT06_17[17]; dat$TT06_22[17]; 
fivecharwords$subject19[9] <- dat$TT06_18[19]#;dat$TT06_17[19]; dat$TT06_18[19]; 
fivecharwords$subject21[9] <- dat$TT06_18[21]#;dat$TT06_17[21]; dat$TT06_18[21]; 
fivecharwords$subject23[9] <- dat$TT06_16[23]#;dat$TT06_17[23]; dat$TT06_16[23]; 
fivecharwords$subject25[9] <- dat$TT06_18[25]#;dat$TT06_17[25]; dat$TT06_18[25]; 
fivecharwords$subject26[9] <- dat$TT06_20[26]#;dat$TT06_17[26]; dat$TT06_20[26]; 
fivecharwords$subject28[9] <- dat$TT06_18[28]#;dat$TT06_17[28]; dat$TT06_18[28]; 

fivecharwords$subject2[10] <- dat$TT06_21[2]#;dat$TT06_20[2]; dat$TT06_21[2]; 
fivecharwords$subject7[10] <- dat$TT06_21[7]#;dat$TT06_20[7]; dat$TT06_21[7]; 
fivecharwords$subject8[10] <- dat$TT06_23[8]#;dat$TT06_20[8]; dat$TT06_23[8]; 
fivecharwords$subject12[10] <- dat$TT06_21[12]#;dat$TT06_20[12]; dat$TT06_21[12]; 
fivecharwords$subject13[10] <- dat$TT06_21[13]#;dat$TT06_20[13]; dat$TT06_21[13]; 
fivecharwords$subject19[10] <- dat$TT06_21[19]#;dat$TT06_20[19]; dat$TT06_21[19]; 
fivecharwords$subject21[10] <- dat$TT06_21[21]#;dat$TT06_20[21]; dat$TT06_21[21]; 
fivecharwords$subject23[10] <- dat$TT06_19[23]#;dat$TT06_20[23]; dat$TT06_19[23]; 
fivecharwords$subject25[10] <- dat$TT06_21[25]#;dat$TT06_20[25]; dat$TT06_21[25]; 
fivecharwords$subject26[10] <- dat$TT06_23[26]#;dat$TT06_20[26]; dat$TT06_23[26]; 
fivecharwords$subject28[10] <- dat$TT06_21[28]#;dat$TT06_20[28]; dat$TT06_21[28]; 
fivecharwords$subject17[10] <- NA

fivecharwords$subject1[11] <- dat$TT06_23[1]#;dat$TT06_22[1]; dat$TT06_23[1]; 
fivecharwords$subject2[11] <- dat$TT06_23[2]#;dat$TT06_22[2]; dat$TT06_23[2]; 
fivecharwords$subject7[11] <- dat$TT06_23[7]#;dat$TT06_22[7]; dat$TT06_23[7]; 
fivecharwords$subject8[11] <- dat$TT06_23[8]#;dat$TT06_22[8]; dat$TT06_23[8]; 
fivecharwords$subject11[11] <- dat$TT06_23[11]#;dat$TT06_22[11]; dat$TT06_23[11]; 
fivecharwords$subject13[11] <- dat$TT06_23[13]#;dat$TT06_22[13]; dat$TT06_23[13]; 
fivecharwords$subject18[11] <- dat$TT06_23[18]#;dat$TT06_22[18]; dat$TT06_23[18]; 
fivecharwords$subject19[11] <- dat$TT06_23[19]#;dat$TT06_22[19]; dat$TT06_23[19]; 
fivecharwords$subject23[11] <- dat$TT06_21[23]#;dat$TT06_22[23]; dat$TT06_21[23]; 
fivecharwords$subject28[11] <- dat$TT06_23[28]#;dat$TT06_22[28]; dat$TT06_23[28]; 
fivecharwords$subject8[11] <- NA
fivecharwords$subject12[11] <- NA
fivecharwords$subject17[11] <- NA
fivecharwords$subject21[11] <- NA
fivecharwords$subject25[11] <- NA
fivecharwords$subject26[11] <- NA

# Necessary updates of word numbers in text 3
fivecharwords$subject16[12] <- dat$TT09_05[16]#;dat$TT09_04[16]; dat$TT09_05[16]; 
fivecharwords$subject27[12] <- dat$TT09_05[27]#;dat$TT09_04[27]; dat$TT09_05[27]; 

fivecharwords$subject7[13] <- dat$TT09_09[7]#;dat$TT09_06[7]; dat$TT09_09[7]; 
fivecharwords$subject16[13] <- dat$TT09_07[16]#;dat$TT09_06[16]; dat$TT09_07[16]; 
fivecharwords$subject27[13] <- dat$TT09_07[27]#;dat$TT09_06[27]; dat$TT09_07[27]; 
fivecharwords$subject28[13] <- dat$TT09_08[28]#;dat$TT09_06[28]; dat$TT09_08[28]; 

fivecharwords$subject7[14] <- dat$TT09_10[7]#;dat$TT09_07[7]; dat$TT09_10[7]; 
fivecharwords$subject16[14] <- dat$TT09_08[16]#;dat$TT09_07[16]; dat$TT09_08[16]; 
fivecharwords$subject27[14] <- dat$TT09_08[27]#;dat$TT09_07[27]; dat$TT09_08[27]; 
fivecharwords$subject28[14] <- dat$TT09_09[28]#;dat$TT09_07[28]; dat$TT09_09[28]; 

fivecharwords$subject7[15] <- dat$TT09_12[7]#;dat$TT09_09[7]; dat$TT09_12[7]; 
fivecharwords$subject11[15] <- dat$TT09_10[11]#;dat$TT09_09[11]; dat$TT09_10[11]; 
fivecharwords$subject16[15] <- dat$TT09_10[16]#;dat$TT09_09[16]; dat$TT09_10[16]; 
fivecharwords$subject27[15] <- dat$TT09_10[27]#;dat$TT09_09[27]; dat$TT09_10[27]; 
fivecharwords$subject28[15] <- dat$TT09_11[28]#;dat$TT09_09[28]; dat$TT09_11[28]; 

fivecharwords$subject4[16] <- dat$TT09_19[4]#;dat$TT09_18[4]; dat$TT09_19[4]; 
fivecharwords$subject5[16] <- dat$TT09_19[5]#;dat$TT09_18[5]; dat$TT09_19[5]; 
fivecharwords$subject7[16] <- dat$TT09_21[7]#;dat$TT09_18[7]; dat$TT09_21[7]; 
fivecharwords$subject11[16] <- dat$TT09_19[11]#;dat$TT09_18[11]; dat$TT09_19[11]; 
fivecharwords$subject13[16] <- dat$TT09_19[13]#;dat$TT09_18[13]; dat$TT09_19[13]; 
fivecharwords$subject16[16] <- dat$TT09_19[16]#;dat$TT09_18[16]; dat$TT09_19[16]; 
fivecharwords$subject17[16] <- dat$TT09_19[17]#;dat$TT09_18[17]; dat$TT09_19[17]; 
fivecharwords$subject18[16] <- dat$TT09_21[18]#;dat$TT09_18[18]; dat$TT09_21[18]; 
fivecharwords$subject21[16] <- dat$TT09_21[21]#;dat$TT09_18[21]; dat$TT09_21[21]; 
fivecharwords$subject22[16] <- dat$TT09_19[22]#;dat$TT09_18[22]; dat$TT09_19[22]; 
fivecharwords$subject26[16] <- dat$TT09_19[26]#;dat$TT09_18[26]; dat$TT09_19[26]; 
fivecharwords$subject27[16] <- dat$TT09_19[27]#;dat$TT09_18[27]; dat$TT09_19[27]; 
fivecharwords$subject28[16] <- dat$TT09_19[28]#;dat$TT09_18[28]; dat$TT09_19[28]; 
fivecharwords$subject30[16] <- dat$TT09_19[30]#;dat$TT09_18[30]; dat$TT09_19[30]; 
fivecharwords$subject12[16] <- NA

fivecharwords$subject1[17] <- dat$TT09_20[1]#;dat$TT09_19[1]; dat$TT09_20[1]; 
fivecharwords$subject4[17] <- dat$TT09_20[4]#;dat$TT09_19[4]; dat$TT09_20[4];
fivecharwords$subject5[17] <- dat$TT09_20[5]#;dat$TT09_19[5]; dat$TT09_20[5]; 
fivecharwords$subject7[17] <- dat$TT09_22[7]#;dat$TT09_19[7]; dat$TT09_22[7]; 
fivecharwords$subject11[17] <- dat$TT09_20[11]#;dat$TT09_19[11]; dat$TT09_20[11]; 
fivecharwords$subject13[17] <- dat$TT09_20[13]#;dat$TT09_19[13]; dat$TT09_20[13]; 
fivecharwords$subject16[17] <- dat$TT09_20[16]#;dat$TT09_19[16]; dat$TT09_20[16]; 
fivecharwords$subject17[17] <- dat$TT09_20[17]#;dat$TT09_19[17]; dat$TT09_20[17]; 
fivecharwords$subject18[17] <- dat$TT09_22[18]#;dat$TT09_19[18]; dat$TT09_22[18]; 
fivecharwords$subject20[17] <- dat$TT09_20[20]#;dat$TT09_19[20]; dat$TT09_20[20]; 
fivecharwords$subject21[17] <- dat$TT09_22[21]#;dat$TT09_19[21]; dat$TT09_22[21]; 
fivecharwords$subject22[17] <- dat$TT09_20[22]#;dat$TT09_19[22]; dat$TT09_20[22]; 
fivecharwords$subject26[17] <- dat$TT09_20[26]#;dat$TT09_19[26]; dat$TT09_20[26]; 
fivecharwords$subject27[17] <- dat$TT09_20[27]#;dat$TT09_19[27]; dat$TT09_20[27]; 
fivecharwords$subject28[17] <- dat$TT09_20[28]#;dat$TT09_19[28]; dat$TT09_20[28]; 
fivecharwords$subject30[17] <- dat$TT09_20[30]#;dat$TT09_19[30]; dat$TT09_20[30]; 
fivecharwords$subject12[17] <- NA

fivecharwords$subject1[18] <- dat$TT09_23[1]#;dat$TT09_22[1]; dat$TT09_23[1]; 
fivecharwords$subject4[18] <- dat$TT09_23[4]#;dat$TT09_22[4]; dat$TT09_23[4]; 
fivecharwords$subject5[18] <- dat$TT09_23[5]#;dat$TT09_22[5]; dat$TT09_23[5]; 
fivecharwords$subject11[18] <- dat$TT09_23[11]#;dat$TT09_22[11]; dat$TT09_23[11]; 
fivecharwords$subject13[18] <- dat$TT09_23[13]#;dat$TT09_22[13]; dat$TT09_23[13]; 
fivecharwords$subject16[18] <- dat$TT09_23[16]#;dat$TT09_22[16]; dat$TT09_23[16]; 
fivecharwords$subject17[18] <- dat$TT09_23[17]#;dat$TT09_22[17]; dat$TT09_23[17]; 
fivecharwords$subject20[18] <- dat$TT09_23[20]#;dat$TT09_22[20]; dat$TT09_23[20]; 
fivecharwords$subject22[18] <- dat$TT09_23[22]#;dat$TT09_22[22]; dat$TT09_23[22];
fivecharwords$subject26[18] <- dat$TT09_23[26]#;dat$TT09_22[26]; dat$TT09_23[26]; 
fivecharwords$subject27[18] <- dat$TT09_23[27]#;dat$TT09_22[27]; dat$TT09_23[27]; 
fivecharwords$subject28[18] <- dat$TT09_23[28]#;dat$TT09_22[28]; dat$TT09_23[28]; 
fivecharwords$subject30[18] <- dat$TT09_23[30]#;dat$TT09_22[30]; dat$TT09_23[30]; 
fivecharwords$subject7[18] <- NA
fivecharwords$subject12[18] <- NA
fivecharwords$subject18[18] <- NA
fivecharwords$subject21[18] <- NA

## split typed word and typing (format so far was "WORD : TIMING")
for(i in 1:nrow(fivecharwords)) {
  for(j in 5:34) {
    #print(fivecharwords[i,j])
    if(!is.na(fivecharwords[i,j])) {
      x <- strsplit(fivecharwords[i,j], " : ")
      word<- x[[1]][1]
      time<- x[[1]][2]
      eval(parse(text=paste0("fivecharwords$subject", j-4,"[", i, "]<- '", 
                             as.character(word), "'")))
      eval(parse(text=paste0("fivecharwords$time_subject", j-4,"[", i, "]<- ",
                             as.numeric(as.character(time)))))
    }
  }
}


## Delete Shift (because it's already applied in the output, e.g. ShiftStuhl) and replace backspace by number
for(i in 1:nrow(fivecharwords)) {
  for(j in 5:34) {
    word <- replace_special_chars(input = fivecharwords[i,j], 
                          oldnames = c("<br>", "Shift", "SHIFT", 
                                       "Backspace","BACKSPACE"), 
                          newnames = c(" ", "", "", "9", "9"))
    eval(parse(text=paste0("fivecharwords$subject", j-4,"[", i, "]<- '", 
                           as.character(word), "'")))
  }}

## Check whether strings were typed correct (if osa > 0, there was some kind of mistake)
for(i in 1:nrow(fivecharwords)) {
  for(j in 5:34) {
    if(fivecharwords[i,j]!="NA") {
      if(fivecharwords[i,j] == toupper(fivecharwords[i,j])) {
        accuracy_wordlevel <- stringdist::stringdist(
        toupper(fivecharwords$word[i]), fivecharwords[i,j], method = "osa", 
          weight = c(d = 1, i = 1, s = 1, t = 1))[1]
        } else {
        accuracy_wordlevel <- stringdist::stringdist(
        fivecharwords$word[i], fivecharwords[i,j], method = "osa", 
          weight = c(d = 1, i = 1, s = 1, t = 1))[1]
        }
        eval(parse(text=paste0("fivecharwords$accuracy_subject", j-4,"[", i, "]<-", 
                             as.numeric(as.character(accuracy_wordlevel)))))
    } else {
      eval(parse(text=paste0("fivecharwords$accuracy_subject", j-4,"[", i, "]<-", 
                           NA)))
    }
}}

## Calculate accuracy per participant
accuracy_fiveletterwords <- NA
for(j in 5:34) {
    eval(parse(text=paste0("incorrect<-sum(fivecharwords$accuracy_subject", j-4,"!=", 0, ", na.rm=T)")))
    eval(parse(text=paste0("total<-sum(!is.na(fivecharwords$accuracy_subject", j-4,"))")))
    # write to accuracy df
    accuracy_fiveletterwords[j-4]<-1-incorrect/total  
}
```

5-character-word based accuracy: 

```{r}
## Descriptive statistics: 
print("Mean Accuracy (5-char words correct)"); round(mean(accuracy_fiveletterwords),2)
print("SD Accuracy (5-char words correct)"); round(sd(accuracy_fiveletterwords),2)
print("Range Accuracy (5-char words correct)"); round(range(accuracy_fiveletterwords),2)
```

Alternative calculation of accuracy: Character-based accuracy on the entire text - Participants typed texts are compared to the entire text and the number of insertions, deletions, substitutions and transpositions are devided by the total amount of characters in the text. 

```{r raw_accuracy}
# delete endings
  dat$TT02_01<- delete_ending(dat$TT02_01)
  dat$TT05_01<-delete_ending(dat$TT05_01)
  dat$TT08_01<-delete_ending(dat$TT08_01)
# delete line break at the end
  dat$TT02_01<-delete_ending(dat$TT02_01, ending = '<br>')
  dat$TT05_01<-delete_ending(dat$TT05_01, ending = '<br>')
  dat$TT08_01<-delete_ending(dat$TT08_01, ending = '<br>')

# replace line break (by space), SHIFT (by nothing) and BACKSPACE  (by nothing)
  # this means that backspace errors are not counted twice, but weighted equally
  # as raw errors. Replace further special keys by numbers so errors are only 
  # counted once
  dat$TT02_01 <- replace_special_chars(input = dat$TT02_01, 
                        oldnames = c("<br>", "Shift", "SHIFT", "Backspace",
                                "BACKSPACE", "CapsLock", "CAPSLOCK", "ArrowRight", 
                                "ArrowUp", "ArrowLeft", "Control", "Delete"), 
                        newnames = c(" ", "", "", "", "", "", "", "1", "1", "1", "1", "1"))
  dat$TT05_01 <- replace_special_chars(input = dat$TT05_01, 
                        oldnames = c("<br>", "Shift", "SHIFT", "Backspace",
                                "BACKSPACE", "CapsLock", "CAPSLOCK", "ArrowRight",
                                "ArrowUp", "ArrowLeft", "Control", "Delete"), 
                        newnames = c(" ", "", "", "", "", "", "", "1", "1", "1", "1", "1"))
  dat$TT08_01 <- replace_special_chars(input = dat$TT08_01, 
                        oldnames = c("<br>", "Shift", "SHIFT", "Backspace",
                                "BACKSPACE", "CapsLock", "CAPSLOCK", "ArrowRight", 
                                "ArrowUp", "ArrowLeft", "Control", "Delete"), 
                        newnames = c(" ", "", "", "", "", "", "", "1", "1", "1", "1", "1"))


# calculate numbers of errors for each subtext
   # osa: insertions, deletions, substitutions and transpositions of
   # adjacent characters are counted and weighted equally
  # participants were instructed to disable caps lock. However, 
  # some did not follow the instructions, here the uppercase solution is 
  # compared to the typed word
accuracy <- data.frame(subject=seq(1,30, by=1),
                          test1_capslock=rep(NA, times=30),
                          test1_raw=rep(NA, times=30),
                          test2_capslock=rep(NA, times=30),
                          test2_raw=rep(NA, times=30),
                          test3_capslock=rep(NA, times=30),
                          test3_raw=rep(NA, times=30))
for (i in 1:nrow(dat)){
  if((dat$TT02_01[dat$subject == i] == toupper(dat$TT02_01[dat$subject == i]) || 
      i == 19 || i == 22)) {
    accuracy$test1_capslock[i] <- "yes"
    # convert : to . and ; to , (this sometimes seems to be a keyboard
    # problem when using caps lock)
    dat$TT02_01[dat$subject == i] <- 
      gsub(":", ".", dat$TT02_01[dat$subject == i], fixed = T)
    dat$TT02_01[dat$subject == i] <- 
      gsub(";", ",", dat$TT02_01[dat$subject == i], fixed = T)
    accuracy$test1_raw[i] <- stringdist::stringdist(
      toupper(test1),dat$TT02_01[dat$subject == i],method = "osa", 
      weight = c(d = 1, i = 1, s = 1, t = 1))[1]
  } else {
    accuracy$test1_capslock[i] <- "no"
    accuracy$test1_raw[i] <- stringdist::stringdist(
      test1, dat$TT02_01[dat$subject == i],method = "osa", 
      weight = c(d = 1, i = 1, s = 1, t = 1))[1]
  }
  
  if(dat$TT05_01[dat$subject == i] == toupper(dat$TT05_01[dat$subject == i])) {
    accuracy$test2_capslock[i] <- "yes"
    # very few participants who had capslock enabled typed '?' instead of 'ß'
    # (this seems to be a keyboard-related problem)
    # replace WEI? by WEIß 
    dat$TT05_01[dat$subject == i] <- 
      gsub("WEI?", "WEIß", dat$TT05_01[dat$subject == i], fixed = T)
    # convert : to . and ; to , (this sometimes seems to be a keyboard
    # problem when using capslock)
    dat$TT05_01[dat$subject == i] <- 
      gsub(":", ".", dat$TT05_01[dat$subject == i], fixed = T)
    dat$TT05_01[dat$subject == i] <- 
      gsub(";", ",", dat$TT05_01[dat$subject == i], fixed = T)
    
    accuracy$test2_raw[i] <- stringdist::stringdist(
      toupper(test2), dat$TT05_01[dat$subject == i], method = "osa", 
      weight = c(d = 1, i = 1, s = 1, t = 1))[1]
  } else {
    accuracy$test2_capslock[i] <- "no"
    accuracy$test2_raw[i] <- stringdist::stringdist(
      test2,  dat$TT05_01[dat$subject == i], method = "osa", 
       weight = c(d = 1, i = 1, s = 1, t = 1))[1]
  }
  
  if((dat$TT08_01[dat$subject == i] == toupper(dat$TT08_01[dat$subject == i]) ||
      i == 14)) {
    accuracy$test3_capslock[i] <- "yes"
    # convert : to . and ; to , (this sometimes seems to be a keyboard
    # problem when using capslock)
    dat$TT08_01[dat$subject == i] <- 
      gsub(":", ".", dat$TT08_01[dat$subject == i], fixed = T)
    dat$TT08_01[dat$subject == i] <- 
      gsub(";", ",", dat$TT08_01[dat$subject == i], fixed = T)
    
    accuracy$test3_raw[i] <- stringdist::stringdist(
      toupper(test3), dat$TT08_01[dat$subject == i],method = "osa", 
      weight = c(d = 1, i = 1, s = 1, t = 1))[1]
  } else {
    accuracy$test3_capslock[i] <- "no"
    accuracy$test3_raw[i] <- stringdist::stringdist(
      test3, dat$TT08_01[dat$subject == i], method = "osa", 
      weight = c(d = 1, i = 1, s = 1, t = 1))[1]
  }
}

# calculate accuracy for each subtext
accuracy$test1_raw <- 1-accuracy$test1_raw/nchar(test1)
accuracy$test2_raw<- 1-accuracy$test2_raw/nchar(test2)
accuracy$test3_raw <- 1-accuracy$test3_raw/nchar(test3)
```

Text-based accuracy: 

```{r text-based accuracy}
accuracy_raw <- (accuracy$test1_raw + accuracy$test2_raw + accuracy$test3_raw)/3
print("Mean Accuracy (entire text)");round(mean(accuracy_raw),2)
print("Sd Accuracy (entire text)");round(sd(accuracy_raw), 2)
print("Range Accuracy (entire text)");round(range(accuracy_raw),2)

cor.test(accuracy_raw, accuracy_fiveletterwords)
```


### Speed
Words per minute: Defined as the number of words correctly typed divided by the total time of all words (similar to Pinet et al., 2016; Crump & Logan 2010)

```{r speed_wpm}
wpm <- NA
for (i in 1:30) {
  eval(parse(text=paste0(
  "x<- fivecharwords %>% filter(!is.na(accuracy_subject",i,"))")))
  # calculate no of words correctly typed
  eval(parse(text=paste0(
  "number_of_words <- sum(x$accuracy_subject", i,"!=", 0, ", na.rm=T)")))
  # calculate total time need to type all fiveletter words
  eval(parse(text=paste0(
  "total_time_fiveletter <- sum(x$time_subject", i,", na.rm=T)")))
  # convert time to seconds
  total_time_fiveletter <- total_time_fiveletter/1000/60
  #calculate word per minutes
  wpm[i] <-number_of_words/total_time_fiveletter 
}
```

Speed in 5-character words per minute:

```{r results speed per min}
## Descriptive statistics: 
print("Mean Speed (5-char words correct)"); round(mean(wpm),2)
print("SD Speed (5-char words correct)"); round(sd(wpm),2)
print("Range Speed (5-char words correct)"); round(range(wpm),2)
```

Alternative speed measurement: Characters per minute, based on the length of the text, not the actual number of characters typed.

```{r speed_totaltext}
speed1 <- nchar(test1)/(dat$TT02_02/100/60)
speed2 <- nchar(test2)/(dat$TT05_02/100/60)
speed3 <- nchar(test3)/(dat$TT08_02/100/60)
```

```{r speed_totaltext_descriptives}
cor.test(speed1, speed2)
cor.test(speed1, speed3)
cor.test(speed2, speed3)
speedperppt <- (speed1+speed2+speed3)/3
cor.test(speedperppt, wpm)
print("mean characters per minute:"); round(mean(speedperppt),2); 
print("SD characters per minute:"); round(sd(speedperppt),2); 
print("range characters per minute:"); round(range(speedperppt),2);
```

## System info (indicated by participants)
Browser: 

```{r browser}
df <- df %>% mutate(browser_char = case_when(browser == 1 ~ "Chrome",
                                             browser == 2 ~ "Coast",
                                             browser == 3 ~ "Firefox",
                                             browser == 4 ~ "Internet Explorer",
                                             browser == 5 ~ "Opera",
                                             browser == 6 ~ "Safari", 
                                             browser == 8 ~ "Microsoft Edge",
                                             browser == 7 ~ browser_other))

table(df$browser_char)
```

Operating system:

```{r OS}
df <- df %>% mutate(operator_system_char = case_when(operator_system == 1 ~ "MacOSX",
                                             operator_system == 2 ~ "Linux",
                                             operator_system == 3 ~ "Windows10",
                                             operator_system == 4 ~ "Windows8",
                                             operator_system == 5 ~ "Windows7",
                                             operator_system == 6 ~ "WindowsVista", 
                                             operator_system == 8 ~ "WindowsNT",
                                             operator_system == -1 ~ "don't know", 
                                             operator_system == -9 ~ "NA",
                                             operator_system == 7 ~ operator_system_other))
table(df$operator_system_char)
```

System:

```{r system}
df <- df %>% mutate(system_char = case_when(system == 1 ~ "Computer(PC)",
                                             system == 2 ~ "Laptop",
                                             system == 3 ~ "TV",
                                             system == 4 ~ "Tablet",
                                             system == 5 ~ "Phone",
                                             system == -1 ~ "don't know", 
                                             system == -9 ~ "NA",
                                             system == 7 ~ "other"))
table(df$system_char)
```



## Attention checks
*1) Item vs. non-item*

```{r item vs non-item}
## Item vs. non-item
# CH01_01 (Taube), CH01_02 (Apfel), CH02_01 (Luftballon) and CH02_02 (Biene) are items and 2 should be selected,
# CH02_03 (Radio), CH02_04 (Sparschwein), CH02_03 (Laptop) and CH02_04 (Wattestäbchen) are non-items and 1 should be selected
## Did participants cheat
# CH03 = 1 - yes, I worked through it till the end,
# CH03 = 2 - no, I stopped or cheated midway
# CH03 = -9 - no answer

attcheck <- data.frame(subject = unique(df$subject)) 

df <- df %>% mutate(itemvsnonitem1 = 
                      case_when(CH01_01==2 & CH01_02==2 & CH01_03==1 & CH01_04==1 ~2,
                                CH01_01==2 || CH01_02==2 ~1, 
                                CH01_01!=2 & CH01_02!=2 ~0)) %>%
  mutate(itemvsnonitem2 = 
                      case_when(CH02_01==2 & CH02_02==2 & CH02_03==1 & CH02_04==1 ~2,
                                CH02_01==2 || CH02_02==2 ~1, 
                                CH02_01!=2 & CH02_02!=2 ~0))
table(df$itemvsnonitem1)/160
table(df$itemvsnonitem2)/160

# attcheck <- data.frame(subject = unique(df$subject))
# 
# 
# data <- data %>% mutate(attcheck =
#                   ifelse(CH01_01 == 2 & CH01_02 == 2 & CH02_01 == 2 & CH02_02 == 2 &
#                   CH01_03 == 1 & CH01_04 == 1 & CH02_03 == 1 & CH02_04 == 1, 1, 0)) %>%
#                    mutate(cheat = ifelse(CH03 == 1,1,ifelse(CH03 == 2,2,0)))
# data.frame(data$subject, )
# table(data$attcheck)
# table(data$cheat)
# 
# # get prolific IDs of participants who failed the attention check
# #pretest %>% subset(attcheck == 0 & cheat == 2 ) %>%
# #  pull(SD24_01) # SD24_01 is prolific ID
# 
# # subset to participants who passed only
# valid <- pretest  %>% filter(attcheck == 1 & cheat != 2)
# 
# inspect <- data.frame(df$subject, df$word, df$fam_typed)

```

(Six persons made one mistake after the familiarization, but all were 100 % after the main experiment, which was the crucial part and thus the relevant criterion)  

*2) Cheating*

```{r cheating}
df <- df %>% mutate(CH03 = case_when(CH03 == 1 ~ 
                                       " Ja, ich habe alles bis zum Ende bearbeitet.",
                                             CH03 == 2 ~ 
                                       "Nein, ich habe zwischendurch aufgehoert oder geschummelt."))
table(df$CH03)/160
```

One person indicated that he/she cheated, but in a comment explained that during the familiarization, he/she took a break to briefly open the door

```{r cheating2}
#table(df$comments[df$CH03 == "Nein, ich habe zwischendurch aufgehoert oder geschummelt."])
```

*3) Keyboard Check*

```{r keyboard}
# self-indicated keyboard type
df <- df %>% mutate(keyboard_type = case_when(keyboard_type == 1 ~ "QWERTY",
                                             keyboard_type == 2 ~ "QWERTZ",
                                             keyboard_type == 3 ~ "QÜERTY",
                                             keyboard_type == 4 ~ "ÄWERTY",
                                             keyboard_type == 5 ~ "AZERTY",
                                             keyboard_type == 6 ~ "QZERTY", 
                                             keyboard_type == 7 ~ "other"))

table(df$keyboard_type)/160

# keyboard test
table(df$KB02_01)/160
table(df$KB03_01)/160 # code ä is Quote
table(df$KB03_02)/160 # code y is KeyZ
table(df$KB03_03)/160 # code p is KeyP
```

(The keyboard screening worked well for all participants)

## Comments
Comments don't indicate any problems that should lead to participant exclusion: 

```{r comments}
#table(df$comments)/160
```



## Fully anonymize data and reduce data frame

```{r anonymize}
df_a <- df %>% dplyr::select(!"gender" & !"age" & !starts_with("language") & 
                               !starts_with("TIME") & !starts_with("handedness") &
                               !starts_with("fingers") & !starts_with("browser") & 
                               !starts_with("operator_system") & !starts_with("system") &
                               !"keyboard_type" & !starts_with("KB") &
                               !starts_with("CH0") & !starts_with("X.") & !"X" &
                               !starts_with("screen") & !starts_with("os") &
                               !starts_with("questionnaire") & !"OR01_01" & 
                               !starts_with("TT0") & !"comments" & !"FINISHED" & !"Q_VIEWER" & 
                               !"LASTPAGE" & !"MAXPAGE" & !"DEG_TIME"& !"type" & !"fam_typed" & 
                               !"gender_char" & !"itemvsnonitem1" & !"itemvsnonitem2")
```

Reduce data frame to columns needed for data analyses or useful to understand the data (this df will be shared online). Interkeystroke intervals could be shared as well, but might only lead to confusion because the data frame will still be very wide.

```{r reduce_df}
df_a <- df_a %>% dplyr::select(c("subject", "trial", "item", "category", "supercategory", "PosOr", 
                        "answercode", "correct", "correct_manual",
                        "answer_auto_jaro", "correct_auto_jaro", 
                        "word", "word.c", "letters.01", "timing.01"))
```

```{r write_anonymized_df}
write.csv(df_a, here::here("data","data_long_anonymous.csv"))
```


