---
title: "05 CSI online typing: Plotting and analysis"
author: "Kirsten Stark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages
```{r load_packages}
library(dplyr)
library(tidyr)
library(lme4)
library(lmerTest)
library(Rmisc)
library(Cairo)
#library(strengejacke)
library(ggplot2)
library(sjPlot)

options(scipen=999)

rm(list = ls())
options( "encoding" = "UTF-8" )
```


## Load and preprocess data

```{r load_data}
# input 
input = "data_long_final.csv"

# load data
df <- read.csv(here::here("data", input))
```

Check amount of participants and trials

```{r checks}
# no. of participants: 
length(unique(df$subject))

# no. of trials is 160 per participant? 
nrow(df) == 160 * length(unique(df$subject))

table(df$answercode)
```

Factorize columns

```{r transform_variables}
# factorize columns
is.numeric(df$timing.01)
df$PosOr <- as.factor(df$PosOr)
df$subject <- as.factor(df$subject)
```

# Plotting

Make plots suitable for APA format, font sizes can be adjusted

```{r apatheme}
apatheme <- theme_bw()+
  theme(plot.title=element_text(family="Arial",size=22,hjust = .5),
        panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        panel.border=element_blank(),axis.line=element_line(),
        text=element_text(family="Arial",size=16))
```

### Descriptives

```{r descriptives}
(means_final<- df %>% filter(!is.na(correct)) %>% 
   filter(category != "Filler") %>% 
   Rmisc::summarySEwithin(.,"timing.01",idvar = "subject",
                          withinvars = "PosOr", na.rm = T))

# Export as word file
library(flextable)
huxt_word <- huxtable::huxtable(means_final)
huxt_word <- huxtable::set_number_format(huxt_word, round(2))
huxtable::quick_docx(huxt_word, 
                     file = here::here("tables",
                                       "CSI_online_typing_RT_summary.docx"), 
                                       open = FALSE)

```

### RTs by ordinal position

```{r plot_rt}
(plot_rt <- df %>% filter(!is.na(correct)) %>% 
   filter(category != "Filler") %>% 
    ggplot(., aes(x=PosOr, y=timing.01, group = correct)) +
    stat_summary(fun=mean,  geom="point", size = 2)+
    stat_summary(fun=mean,  geom="line", size = 1) +
    apatheme+
    labs(x="Ordinal Position ",y ="RT (ms)")+
   annotate(geom="text", x=1.5, y=1330, label="n = 30", 
           color="black", size = 8))


filename <- "CSI_online_typing_plot_rt.pdf"
ggsave(plot_rt, filename = 
         here::here("figures", filename),
       width = 18, height = 13, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("figures", filename))

```

### ... with fillers for control

```{r plot_rt_fillers}

(plot_rt_fillers <- df %>% filter(!is.na(correct)) %>% 
    mutate(kind = case_when(category == "Filler" ~"Filler",
                          category != "Filler" ~"Experimental")) %>%
    ggplot(., aes(x=PosOr, y=timing.01, group=kind, color=kind)) +
    stat_summary(fun=mean,  geom="point", size = 2)+
    stat_summary(fun=mean,  geom="line", size = 1) +
    apatheme+
    labs(x="Ordinal Position ",y ="RT (ms)", color = "Trial type")+
  annotate(geom="text", x=1.5, y=1350, label="n = 30", 
           color="black", size = 8))

filename <- "CSI_online_typing_plot_rt_with_fillers.pdf"
ggsave(plot_rt_fillers, filename = 
         here::here("figures", filename),
       width = 18, height = 13, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("figures", filename))
```
  
### Plot by subcategory

```{r plot_rt_by_cat}
(plot_rt_by_cat <- df %>% filter(!is.na(correct)) %>% 
   filter(category != "Filler") %>% 
    ggplot(., aes(x=PosOr, y=timing.01)) +
    stat_summary(fun=mean,  geom="point", size = 2)+
    stat_summary(fun=mean,  geom="line", size = 1) +
    facet_wrap(~category) +
    apatheme+
    labs(x="Ordinal Position ",y ="RT (ms)"))

filename <- "CSI_online_typing_plot_rt_by_category.pdf"
ggsave(plot_rt_by_cat, filename = 
         here::here("figures", filename),
       width = 18, height = 19, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("figures", filename))

```

### Plot by subject

```{r plot_by_subject}
(plot_rt_by_subject <- df %>% filter(!is.na(correct)) %>% 
   filter(category != "Filler") %>% 
    ggplot(., aes(x=PosOr, y=timing.01)) +
    stat_summary(fun=mean,  geom="point", size = 2) +  
    stat_summary(fun=mean,  geom="line", size = 1) +  
    facet_wrap(~subject) +
    apatheme+
    labs(x="Ordinal Position ",y ="RT (ms)"))

filename <- "CSI_online_typing_plot_rt_by_subject.pdf"
ggsave(plot_rt_by_subject, filename = 
         here::here("figures", filename),
       width = 18, height = 19, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("figures", filename))

```

### Control: Plot RTs accross the experiment
All trials including uncorrect trials

```{r plot_RTs_all}
(plot_RTs_all <- ggplot(data=df, aes(x=trial, y=timing.01)) +
  stat_summary(fun=mean,  geom="point", size = 2)+
  stat_summary(fun=mean,  geom="line", size = 1) +
  apatheme+
  labs(x="Trial ",y ="RT (ms)")+
  annotate(geom="text", x=20, y=1570, label="n = 30", 
           color="black", size = 8))

filename <- "CSI_online_typing_plot_rts_across_experiment.pdf"
ggsave(plot_RTs_all, filename = 
         here::here("figures", filename),
       width = 18, height = 13, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("figures", filename))
```

Correct non-filler trials only:

```{r plot_RTs_correct}
(plot_RTs_correct <- df %>% filter(!is.na(correct)) %>% 
   filter(category != "Filler") %>% 
    ggplot(., aes(x=trial, y=timing.01)) +
    stat_summary(fun=mean,  geom="point", size = 2)+
    stat_summary(fun=mean,  geom="line", size = 1) +
    apatheme+
    labs(x="Trial ",y ="RT ms")+
    annotate(geom="text", x=20, y=1570, label="n = 30", 
             color="black", size = 8))

filename <- "CSI_online_typing_plot_rts_across_experiment_correct_experimental_trials.pdf"
ggsave(plot_RTs_correct, filename = 
         here::here("figures", filename),
       width = 18, height = 13, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("figures", filename))
```

# Check distribution of data
Are the data normally distributed or does a gamma distribution fit the data better?  
*Subset data to correct trials only and excude fillers*

```{r final_data}
df_valid <- df %>% filter(category != "Filler") %>% 
  filter(correct == 1) %>% droplevels()
```

*Center predictor variable*

```{r}
df_valid$PosOr.cont <- scale(as.numeric(as.character(df_valid$PosOr)),
                                        center = T, scale = F)
# table(df_valid$PosOr.cont)
# mean(df_valid$PosOr.cont)
```

*Histogram of the reaction time data*

```{r RT_hist}
hist(df_valid$timing.01)
```

*Check fit of normal vs gamma distribution in histograms, q-q-plots and using objective criteria:*  
1) Fit normal and gamma distributions to the reaction time data

```{r load_fitdistrplus}
library(fitdistrplus)
```

```{r fit.normal}
fit.normal<- fitdist(df_valid$timing.01, distr = "norm", method = "mle")
summary(fit.normal)
#plot(fit.normal)
```

```{r fit.gamma}
fit.gamma <- fitdist(df_valid$timing.01, distr = "gamma", method = "mle")
summary(fit.gamma)
#plot(fit.gamma)
```

2) Compare the fit of the two distributions  
Visually compare fit of both distributions in histogram

```{r density-comparison}
denscomp(list(fit.gamma, fit.normal))
```
Visually compare fit of both distributions in Q-Q-plots

```{r q-q-comparison}
qqcomp(list(fit.gamma, fit.normal))
```

Compare information criteria

```{r fit_criteria-comparison}
gofstat(list(fit.gamma, fit.normal),
        fitnames = c("Gamma", "Normal"))
```

**Conclusion:** Both the visual inspection and the objective criteria suggest that a gamma distribution fits the data better (although not that well). Therefore, we fit a Gamma distribution in a GLMM with the continuous predictor ordinal position (Pos.cont), the factorial predictor (experiment), and their interaction. We compute the maximal random effects structure.


# Inferential analyses: GLMM (Gamma distribution) with ordinal position as a continuous predictor


```{r GLMM_cont}
m1 <- glmer(timing.01 ~ PosOr.cont + 
               (PosOr.cont|subject) +(PosOr.cont|category),
             data = df_valid, 
            family =Gamma(link ="identity"), 
            control=glmerControl(optimizer = "bobyqa"))
summary(m1)

# save model output
tab_model(m1,transform = NULL,
          show.re.var = F, show.stat = T,show.r2 = F,show.icc = F,
          title = "GLMM (Gamma distribution) with continuous predictor",
          pred.labels = c("(Intercept)", "Ordinal Position"),
          dv.labels = "Typing Onset Latency",
          #string.pred = "",
          string.stat = "t-Value",
          file = here::here("tables", "CSI_online_typing_glmm_cont.html"))
```

# ---------------------------------------------
# Other
# Plotting before data preprocessing
# ALL TRIALS
### Plot RTs by OrPos

```{r}
# is.numeric(df$timing.01)
# df$PosOr <- as.factor(df$PosOr)
# df$subject <- as.factor(df$subject)
# # descriptives 
# (means <- Rmisc::summarySEwithin(df[df$category != "Filler" ,],"timing.01",idvar = "subject",withinvars = "PosOr", na.rm = T))
```


```{r}
# plotting <- df %>%
#   mutate(kind = case_when(category == "Filler" ~"Filler",
#                           category != "Filler" ~"Experimental"))
# (plot <- ggplot(data=df[df$category != "Filler" ,], aes(x=PosOr, y=timing.01)) +
#   stat_summary(fun=mean,  geom="point", size = 2)+
#   stat_summary(fun=mean,  geom="line", size = 1) +
#   apatheme+
#   labs(x="Ordinal Position ",y ="RT ms")+
#   annotate(geom="text", x=2, y=1200, label="n = 4", color="black", size = 8))
# 
# (plot <- ggplot(data=plotting, aes(x=PosOr, y=timing.01, group = kind, color = kind)) +
#   stat_summary(fun=mean,  geom="point", size = 2)+
#   stat_summary(fun=mean,  geom="line", size = 1) +
#   apatheme+
#   labs(x="Ordinal Position ",y ="RT ms", color = "Trial type")+
#   annotate(geom="text", x=2, y=1380, label="n = 30", color="black", size = 8))
# 
# 
# # ggsave(plot = plot, file = "figures/RT_across_conditions.pdf", device = cairo_pdf, dpi = 300)
# # embedFonts(file = "figures/RT_across_conditions_CSI_online_typing.pdf")
```
  
Plot by subcategory
```{r}

# (plot_by_cat <- ggplot(data=df[df$category != "Filler" ,], aes(x=PosOr, y=timing.01)) +
#   stat_summary(fun=mean,  geom="point", size = 2)+
#   stat_summary(fun=mean,  geom="line", size = 1) +
#   facet_wrap(~category) +
#   apatheme+
#   labs(x="Ordinal Position ",y ="RT ms"))
# #ggsave(plot = plot30_facet, file = "Plots/CSI_subcats.pdf", device = cairo_pdf,dpi = 300)
# #embedFonts(file = "Plots/CSI_online30_subcats.pdf")

```

Plot by subject: 
```{r}
# 
# (plot_by_cat <- ggplot(data=df[df$category != "Filler" ,], aes(x=PosOr, y=timing.01)) +
#   stat_summary(fun=mean,  geom="point", size = 2)+
#   stat_summary(fun=mean,  geom="line", size = 1) +
#   facet_wrap(~subject) +
#   apatheme+
#   labs(x="Ordinal Position ",y ="RT ms"))
# #ggsave(plot = plot30_facet, file = "Plots/CSI_subcats.pdf", device = cairo_pdf,dpi = 300)
# #embedFonts(file = "Plots/CSI_online30_subcats.pdf")

```

# Other

Full GLMM with polynomial contrasts

```{r}
# contrasts(df$PosOr) <- contr.poly(5)
# # m1 <- glmer(timing.01 ~ PosOr + (PosOr|subject) +(PosOr|category) ,
# #              data = df[df$category != "Filler"  & df$correct == 1,], family =Gamma(link ="identity"), control=glmerControl(optimizer = "bobyqa"))
# #model does not converge
# m1 <- afex::lmer_alt(timing.01 ~ PosOr + (PosOr||subject) +(1|category), 
#                   df[df$category != "Filler"  & df$correct == 1,],
#                   family =Gamma(link ="identity"),
#                   control=glmerControl(optimizer = "bobyqa", 
#                                        optCtrl = list(maxfun = 2*10^5)))
# summary(m1)
```

Linear trend only

````{r}
# contrasts(df$PosOr) <- contr.poly(5)
# fixef_terms <- model.matrix( ~ PosOr,df) 
# df$PosOr.L <- fixef_terms[,2]
# m2 <- glmer(timing.01 ~ PosOr.L + (PosOr.L|subject) +(PosOr.L|category) ,
#              data = df[df$category != "Filler"  & df$correct == 1,], 
#             family =Gamma(link ="identity"), 
#             control=glmerControl(optimizer = "bobyqa"))
# summary(m2)
```

Log-transformed data

```{r LMM_log_cont}
# m1.log <- lmer(log(timing.01) ~ PosOr.cont + 
#                (PosOr.cont|subject) +(1|category),
#              data = df_valid,  
#             control=lmerControl(optimizer = "bobyqa", 
#                                 optCtrl = list(maxfun = 2*10^5)))
# summary(m1.log)
# 
# # save model output
# tab_model(m1.log,transform = NULL,
#           show.re.var = F, show.stat = T,show.r2 = F,show.icc = F,
#           title = "GLMM (Gamma distribution) with continuous predictor",
#           pred.labels = c("(Intercept)", "Ordinal Position"),
#           dv.labels = "Typing Onset Latency",
#           #string.pred = "",
#           string.stat = "t-Value",
#           file = here::here("tables", "CSI_online_typing_lmm_log_cont.html"))
```

```{r}
# anova(m1, m1.log)
```
