---
title: "06 CSI online typing: Comparison to verbal naming"
author: "Kirsten Stark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r load packages}
library(dplyr)
library(tidyr)
library(lme4)
library(lmerTest)
library(Rmisc)
library(Cairo)
#library(strengejacke)
library(ggplot2)

options(scipen=999)

rm(list = ls())
options( "encoding" = "UTF-8" )
```

# Load data
Load data from both the verbal online CSI experiment and the typing online CSI experiment.  
  
Load typing data

```{r load_typing_data}
df_typing <- read.csv(here::here("data", "data_long_final.csv"))
```

Load verbal data

```{r load_verbal_data}
load(here::here("data", "verbal_CSI", "CSI_online_verbal_df_full.RData"))
df_verbal <- df_full
```

# Combine both data frames into one
1) Subset relevant columns and give identical names

```{r subset_and_rename_columns}
df_typing <- df_typing %>% 
  select(subject, item, category, timing.01, PosOr, correct) %>%
  dplyr::rename(RT = timing.01, Pos = PosOr) %>%
  mutate(experiment = "typing")

df_verbal <- df_verbal %>%
  select(VP, Item, subcat, VOT, correct, Pos) %>%
  dplyr::rename(subject = VP, item = Item, category = subcat, RT = VOT) %>%
  mutate(experiment = "verbal")
```

2) Give subjects from both experiments different names

```{r adapt_subject_names}
df_typing <- df_typing %>% mutate(subject = subject + 200)
df_verbal <- df_verbal %>% mutate(subject = subject + 100)
```


3) Put columns into correct format 

```{r factorize}
df_typing <- df_typing %>% 
  mutate(subject = as.factor(subject)) %>%
  mutate(item = as.character(item)) %>%
  mutate(category = as.factor(category)) %>% 
  mutate(RT = as.numeric(RT)) %>%
  mutate(Pos = as.factor(Pos)) %>% 
  mutate(experiment = factor(experiment, levels = c("verbal", "typing")))

df_verbal <- df_verbal %>% 
  mutate(subject = as.factor(subject)) %>%
  mutate(item = as.character(item)) %>%
  mutate(category = as.factor(category)) %>% 
  mutate(RT = as.numeric(RT)) %>%
  mutate(Pos = as.factor(Pos)) %>% 
  mutate(experiment = factor(experiment, levels = c("verbal", "typing")))
```

4) Bind both data frames into one

```{r combine_df}
df <- bind_rows(df_typing, df_verbal)
```

5) Give identical category names in both experiments

```{r category_names}
df <- df %>% mutate(category = case_when(category == "Buero" ~ "Büro",
                                  category == "Gebaeude" ~ "Gebäude",
                                  category == "Gemuese" ~ "Gemüse",
                                  category == 
                                    "Koerperteile" ~ "Körperteile",
                                  category == "Kueche" ~ "Küche",
                                  category == 
                                    "Suessigkeiten" ~ "Süssigkeiten",
                                  category == 
                                    "Trinkgefaesse" ~ "Trinkgefässe",
                                  category == "Voegel" ~ "Vögel",
                                  TRUE ~ as.character(category))) %>%
  mutate(category == as.factor(category)) %>% droplevels()
table(df$category)
```

5) Drop filler trials

```{r drop_filler}
 df <- df %>% filter(category != "Filler" & 
                      category != "Filler1" & category != "Filler2") %>%
  droplevels()
```


5) Add ordinal position as continuous predictor variable

```{r Pos_cont}
df$Pos.cont <- scale(as.numeric(as.character(df$Pos)),
                       center = T, scale = F)
```

3) Drop incorrect trials

```{r subset_to_correct}
df <- df %>% filter(!is.na(correct) & correct != 0) %>% 
  select(-correct) %>%
  droplevels()
```


# Descriptives

```{r descriptives}
(descriptives <- df %>% 
   Rmisc::summarySEwithin(.,"RT",idvar = "subject",
                          withinvars = "Pos",
                          betweenvars = "experiment",
                          na.rm = T))
```

# Plotting
Make plots suitable for APA format

```{r apatheme}
apatheme <- theme_bw()+
  theme(plot.title=element_text(family="Arial",size=22,hjust = .5),
        panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        panel.border=element_blank(),axis.line=element_line(),
        text=element_text(family="Arial",size=16))
```

Plot RTs by ordinal position for both experiments

```{r plot}
(plot <- df %>% 
    ggplot(., aes(x=Pos, y=RT, 
                  group=experiment, color=experiment)) +
    stat_summary(fun=mean,  geom="point", size = 2)+
    stat_summary(fun=mean,  geom="line", size = 1) +
    apatheme+
    labs(x="Ordinal Position ",y ="RT (ms)", color = "Experiment"))

filename <- "CSI_online_exp_comparison_plot_rt.pdf"
ggsave(plot, filename = 
         here::here("figures", filename),
       width = 18, height = 13, units = "cm", 
       dpi = 300, device = cairo_pdf)
embedFonts(file = here::here("figures", filename))
```

# GLMMs
Check distribution

```{r RT_hist}
hist(df$RT)
```

Set contrasts

```{r set contrasts}
# define sliding difference contrast for factor experiment: 
# contrast is verbal - typing, intercept being the grand mean 
levels(df$experiment)
contrasts(df$experiment) <- MASS::contr.sdif(2)

# define polynomial contrasts for factor Position
contrasts(df$Pos) <- contr.poly(5)
# select linear trend
fixef_terms <- model.matrix( ~ Pos,df) 
df$Pos.L <- fixef_terms[,2]
```

### Full GLMM with polynomial contrasts (linear trend only)

```{r glmm_poly_lin}

# define model
m1 <- glmer(RT ~ Pos.L*experiment +
                       (Pos.L|subject) + (Pos.L*experiment|category),
            data = df,
            family =Gamma(link ="identity"),
            control=glmerControl(optimizer = "bobyqa"))
summary(m1)
```

### Full GLMM with polynomial contrasts (all contrasts included):

```{r glmm_poly}
m2 <- afex::lmer_alt(RT ~ Pos*experiment +
                       (Pos||subject) + (Pos*experiment||category),
            data = df,
            family =Gamma(link ="identity"),
            control=glmerControl(optimizer = "bobyqa",
                                   optCtrl = list(maxfun = 2*10^5)))
summary(m2)
```

### Ordinal position as a continuous predictor variable

```{r glmm_cont}
m4 <- glmer(RT ~ Pos.cont*experiment +
               (Pos.cont|subject) +(Pos.cont*experiment|category),
             data = df,
            family =Gamma(link ="identity"),
            control=glmerControl(optimizer = "bobyqa"))
summary(m4)
```

